manuscript_outline

### 2025-07-14 docs/manuscript/paper.Rmd
table spacing
removed some verbiage from paper

Examining variation in development across space and time is critical to understand how organisms respond to ecological factors such as climate change [@pinsky2019]. Here, we first demonstrate an application of hatchR to predict hatching times from a spatio-temporally extensive dataset. Using these predictions, we then show how one can examine putative drivers of variation in development and discuss their application in forecasting future responses.

@@

This example illustrates how **hatchR** can be used to explore the intersection of developmental plasticity and local adaptation using real-world data. Such applications are useful for disentangling the genetic basis of thermal sensitivity, anticipating geographic responses to climate change, and informing population-specific management strategies.

@@

Although initially developed for fishes, **hatchR** is a broadly applicable tool for modeling temperature-dependent development in poikilotherms. It addresses key limitations of traditional degree-day approaches by accounting for non-linear thermal response curves and enabling predictions under fluctuating thermal conditions. As global temperatures rise and the phenology of organisms shifts in response, tools like **hatchR** will be increasingly valuable for anticipating the ecological and evolutionary consequences of climate change. While we provide numerous species-level parameterizations and demonstrate a set of representative case studies, the full range of possible applications—across taxa, ecosystems, and educational settings—extends far beyond what we present here.

>>>>

Examining variation in development across space and time is critical to understand how organisms respond to ecological factors such as changing environments [@pinsky2019]. Here, we first demonstrate an application of hatchR to predict hatching times from a spatio-temporally extensive dataset. Using these predictions, we then show how one can examine putative drivers of variation in development and discuss their application in forecasting future responses.

@@

This example illustrates how **hatchR** can be used to explore the intersection of developmental plasticity and local adaptation using real-world data. Such applications are useful for disentangling the genetic basis of thermal sensitivity, anticipating geographic responses to changing environments, and informing population-specific management strategies.

@@

Although initially developed for fishes, **hatchR** is a broadly applicable tool for modeling temperature-dependent development in poikilotherms. It addresses key limitations of traditional degree-day approaches by accounting for non-linear thermal response curves and enabling predictions under fluctuating thermal conditions. As global temperatures rise and the phenology of organisms shifts in response, tools like **hatchR** will be increasingly valuable for anticipating the ecological and evolutionary consequences of variable climates. While we provide numerous species-level parameterizations and demonstrate a set of representative case studies, the full range of possible applications—across taxa, ecosystems, and educational settings—extends far beyond what we present here.
